---
title: "christina_notes"
author: "Christina Chen"
date: "2024-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(janitor)
library(naniar)
```

```{r read data}
spacemission <- read_csv("data/space_missions.csv") %>% clean_names()
```

```{r glimpse}
glimpse(spacemission)
```

```{r percentage of NAs}
miss_var_summary(spacemission)
```

```{r plotting launches by year}
spacemission %>%
  separate(date, into = c("year", "month", "date"), sep = "-") %>%
  group_by(year) %>%
  ggplot(aes(x = year)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### notable events
- space race was around:          1951 - 1993
- kennedy speech for man on moon: 1961
- moon landing                    1969

### stuff i had to fix by editing the csv
- added "USA" to new mexico (x3)
- fixed spelling of "plateform" to "platfrom (x1)
- fixed special characters of "Alacantara, Maranhao, Brazil" (x3)
- added "Kekaha, Hawaii, USA" to Pacific Missile Range Facility (x1)
- added "Spain" to Gran Canaria (x2)
- added "International Waters" to Pacific Ocean (x36)
- added "International Waters" to Barents Sea (x3)
- added "International Waters" to Yellow Sea (x3)

```{r reading fixed data}
spacemission_fixed <- read_csv("data/space_missions_fixed.csv") %>% clean_names()
```

```{r separates location and picks out state and country}
spacemission_fixed %>%
  separate(date, into = c("year", "month", "date"), sep = "-") %>%
  separate(location, into = c("site1", "site2", "site3", "site4", "state", "country"), extra = "drop", fill = "left", sep = ",") %>%
  select(-site1, -site2, -site3, -site4)
```

```{r plotting number of space launches by country}
spacemission_fixed %>%
  separate(location, into = c("site1", "site2", "site3", "site4", "state", "country"), extra = "drop", fill = "left", sep = ",") %>%
  select(-site1, -site2, -site3, -site4) %>%
  group_by(country) %>%
  summarise(total = n()) %>%
  ggplot(aes(x = reorder(country, total), y = total, fill = country)) +
  geom_col(show.legend = F) + 
  coord_flip() +
  geom_text(aes(label = total), hjust = -0.2, size = 2, color = "black") + # the first blank column is Brazil
  theme_classic() + 
  labs(title = "Space Missions by Country",
       x = "Country",
       y = "Total Missions") +
  theme(plot.title = element_text(hjust = 0.5))
```

