```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(janitor)
library(naniar)
```

```{r}
spacemission <- read_csv("data/space_missions.csv") %>% clean_names()
```

```{r}
spacemission %>%
  summarize(mean_price = mean(price, na.rm = T),
            max_price = max(price, na.rm = T),
            min_price = min(price, na.rm = T))
```
Most expensive- 5,000,000 from USSR
```{r}
spacemission %>%
  arrange(desc(price))
```

Least expensive- 2,500,000 from Pacific Spaceport Complex, Kodiak, Alaska
```{r}
spacemission %>%
  arrange(price)
```

```{r}
spacemission_fixed <- read_csv("data/space_missions_fixed.csv") %>% clean_names()
```

## How much each country has spent on rockets on average / mission
```{r}
spacemission_dates <- spacemission %>% clean_names() %>%
  separate(date, into = c("year", "month", "date"), sep = "-")
```


```{r}
spacemission_fixed <- spacemission_fixed %>%
  separate(location, into = c("site1", "site2", "site3", "site4", "state", "country"), extra = "drop", fill = "left", sep = ",") %>%
  select(-site1, -site2, -site3, -site4)
```

## Mean cost of missions by country
```{r}
spacemission_fixed %>%
  group_by(country) %>%
  summarize(mean_cost_by_country = mean(price, na.rm = T)) %>%
  filter(!is.na(mean_cost_by_country)) %>%
  arrange(desc(mean_cost_by_country))
```
## Proportion of failures to successes
```{r}
successful_missions <- spacemission_attempt %>%
  group_by(country) %>%
  filter(mission_status == "Success") %>%
  count(country) %>%
  mutate(number_of_sucesses = n) %>%
  arrange(desc(n))
successful_missions
```

```{r}
failed_missions <- spacemission_attempt %>%
  group_by(country) %>%
  filter(mission_status == "Failure") %>%
  count(country) %>%
  mutate(number_of_failures = n) %>%
  arrange(desc(n))
failed_missions
```
## Number of Successes Versus Number of Failures by Country
```{r}
spacemission_fixed %>%
  separate(date, into = c("year", "month", "date"), sep = "-") %>%
  group_by(country, mission_status) %>%
  mutate(mission_status = as.factor(mission_status)) %>%
  ggplot(aes(x=country, y=mission_status, fill=mission_status))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 90, vjust = .5))+
  theme(legend.position = "bottom")+
    labs(title = "Proportion of Mission Status By Country Between 1957-2022",
       x = "Country",
       y = "Mission Status")
```



## Number of Successes Versus Number of Failures
```{r}
spacemission_dates %>%
  group_by(year, mission_status) %>%
  ggplot(aes(x=year, fill = mission_status))+
  geom_bar(na.rm = T)+
  theme(axis.text.x = element_text(angle = 90, vjust = .5))+
  theme(legend.position = "bottom")+
  labs(title = "Mission Status Between 1957-2022",
       y = "Year",
       x = "Mission Status")
```


## Number of active rockets over time
```{r}
spacemission_dates %>%
  group_by(year, rocket_status) %>%
  ggplot(aes(x=year, fill = rocket_status))+
  geom_bar(na.rm = T)+
  theme(axis.text.x = element_text(angle = 90, vjust = .5))+
  theme(legend.position = "bottom")+
  labs(title = "Proportion of Active Rockets to Inactive Rockets Between 1957-2022",
       x = "Year",
       y = "Number of Rockets")
```

# Unique rockets
```{r}
spacemission %>%
  group_by(rocket)
```

